dispRity palaeo demo

### 2016-05-29 doc/dispRity-palaeo-demo.Rmd
Added code to the Rmd docs (removed redundant Rnw)

\

\
\
\
\
\

    ##                    [,1]          [,2]        [,3]       [,4]      [,5]
    ## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
    ## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
    ## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
    ## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
    ## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
    ## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273

    ## [1] 50 48

\
\
\
\

    ##             FAD  LAD
    ## Adapis     37.2 36.8
    ## Asioryctes 83.6 72.1
    ## Leptictis  33.9 33.3
    ## Miacis     49.0 46.7
    ## Mimotona   61.6 59.2
    ## Notharctus 50.2 47.0

\
\
`  `

@@

` `\
\
`  `

@@

\
`          `\
\
\
`     `\
`           `\
`     `

@@

\

    ## [1] "dispRity"

\

    ## List of 3
    ##  $ data    :List of 5
    ##   ..$ 100-80: num [1:8, 1:48] -0.739 -0.68 -0.511 -0.473 -0.459 ...
    ##   .. ..- attr(*, "dimnames")=List of 2
    ##   .. .. ..$ : chr [1:8] "Daulestes" "Bulaklestes" "Uchkudukodon" "Asioryctes" ...
    ##   .. .. ..$ : NULL
    ##   ..$ 80-60 : num [1:15, 1:48] -0.477 -0.473 -0.474 -0.532 -0.409 ...
    ##   .. ..- attr(*, "dimnames")=List of 2
    ##   .. .. ..$ : chr [1:15] "Kennalestes" "Asioryctes" "Ukhaatherium" "Cimolestes" ...
    ##   .. .. ..$ : NULL
    ##   ..$ 60-40 : num [1:13, 1:48] -0.165 0.254 0.622 0.753 0.671 ...
    ##   .. ..- attr(*, "dimnames")=List of 2
    ##   .. .. ..$ : chr [1:13] "Todralestes" "Pezosiren" "Tribosphenomys" "Paramys" ...
    ##   .. .. ..$ : NULL
    ##   ..$ 40-20 : num [1:6, 1:48] -0.121 -0.203 -0.158 0.696 0.355 ...
    ##   .. ..- attr(*, "dimnames")=List of 2
    ##   .. .. ..$ : chr [1:6] "Leptictis" "Dilambdogale" "Widanelfarasia" "Moeritherium" ...
    ##   .. .. ..$ : NULL
    ##   ..$ 20-0  : num [1:10, 1:48] 0.3079 0.6531 0.5089 -0.0419 0.235 ...
    ##   .. ..- attr(*, "dimnames")=List of 2
    ##   .. .. ..$ : chr [1:10] "Dasypodidae" "Bradypus" "Myrmecophagidae" "Potamogalinae" ...
    ##   .. .. ..$ : NULL
    ##  $ elements: chr [1:50] "Cimolestes" "Maelestes" "Batodon" "Bulaklestes" ...
    ##  $ series  : chr [1:6] "discrete" "100-80" "80-60" "60-40" ...
    ##  - attr(*, "class")= chr "dispRity"

    ## [1] "data"     "elements" "series"

\

    ## 5 discrete series for 50 elements 
    ## Series:
    ## 100-80, 80-60, 60-40, 40-20, 20-0.

@@

\
`     `\
\
\

    ## [1] 6

\
`     `\
`      `\
\
\
` `

    ## [1] "dispRity"
    ## [1] "dispRity"

\

    ## Bootstrapped ordinated matrix with 50 elements 
    ## Series:
    ## 100-80, 80-60, 60-40, 40-20, 20-0.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

    ## Bootstrapped ordinated matrix with 50 elements 
    ## Series:
    ## 100-80, 80-60, 60-40, 40-20, 20-0.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.
    ## Data was rarefied with a maximum of 6 elements

Calculating disparity
---------------------

@@

\
`     `\
\
\
`     `\
\
\
\

    ## Disparity measurements across 5 series for 50 elements 
    ## Series:
    ## 100-80, 80-60, 60-40, 40-20, 20-0.
    ## Disparity calculated as: c(sum, variances) for 48 dimensions.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

    ## Disparity measurements across 5 series for 50 elements 
    ## Series:
    ## 100-80, 80-60, 60-40, 40-20, 20-0.
    ## Disparity calculated as: c(sum, variances) for 48 dimensions.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.
    ## Data was rarefied with a maximum of 6 elements

\

    ##   series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1 100-80  8    1.675 1.460 1.087 1.389 1.568 1.648
    ## 2  80-60 15    1.782 1.678 1.538 1.631 1.728 1.792
    ## 3  60-40 13    1.913 1.773 1.607 1.734 1.826 1.886
    ## 4  40-20  6    2.022 1.673 1.212 1.537 1.822 1.942
    ## 5   20-0 10    1.971 1.773 1.598 1.716 1.842 1.890

    ##   series n observed  mean  2.5%   25%   75% 97.5%
    ## 1 100-80 6       NA 1.440 1.067 1.339 1.575 1.726
    ## 2  80-60 6       NA 1.670 1.276 1.614 1.798 1.922
    ## 3  60-40 6       NA 1.763 1.458 1.704 1.846 1.986
    ## 4  40-20 6     2.02 1.721 1.392 1.633 1.828 1.992
    ## 5   20-0 6       NA 1.762 1.459 1.642 1.880 2.009

@@

\

@@

\
`          `\
\
\
`      `\
`      `

@@

\
`     `\
`      `

    ##                   W      p.value
    ## 100-80 - 80-60  471 7.427563e-28
    ## 80-60 - 60-40  1562 1.798899e-16
    ## 60-40 - 40-20  6250 9.061511e-03
    ## 40-20 - 20-0   3725 7.379715e-03

\
`     `\
`      `

    ##                   W      p.value
    ## 100-80 - 80-60 1695 2.717026e-15
    ## 80-60 - 60-40  3424 4.733196e-04
    ## 60-40 - 40-20  5669 4.095267e-01
    ## 40-20 - 20-0   4254 2.740887e-01

>>>>

```{r}
## Loading demo and the package data
library(dispRity)

## Setting the random seed for repeatability
set.seed(123)

## The morphospace:
data(BeckLee_mat50)
head(BeckLee_mat50[,1:5])

dim(BeckLee_mat50)
## The morphospace contains 50 taxa and has 48 dimensions (or axis)

## A list of first and last occurrences data for some fossils
data(BeckLee_ages)
head(BeckLee_ages)

## And a phylogeny:
data(BeckLee_tree)
plot(BeckLee_tree, cex=0.7); axisPhylo(root=140)
```

@@

```{r, eval = FALSE}
if(!require(strap)) install.packages("strap")
library(strap)
geoscalePhylo(BeckLee_tree, cex.tip=0.7, cex.ts=0.6)
```

@@

```{r}
## Creating the vector of time bins ages
time_bins <- rev(seq(from = 0, to = 100, by = 20))

## Splitting the morphospace using the time.series function
binned_morphospace <- time.series(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages)
```

@@

```{r}
## The object class
class(binned_morphospace)

## What's in the object (messy version for the console!)
str(binned_morphospace)
names(binned_morphospace)

## How it's printed by default
binned_morphospace
```

@@

```{r}
## Bootstrapping each time series 100 times
boot_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100)

## Getting the minimum number of rows (=taxa) in the time seriess
min(unlist(lapply(binned_morphospace$data, nrow)))

## Bootstrapping each time series 100 times and rarefying them 
rare_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100,
    rarefaction = 6)

## Note that the output objects are both from the "dispRity" class ...
class(boot_bin_morphospace); class(rare_bin_morphospace)

## ... and display some more information on the previous step when printed:
boot_bin_morphospace
rare_bin_morphospace
```
Calculating disparity
---------------------

@@

```{r}
## Disparity for the bootstrapped data
boot_disparity <- dispRity(boot_bin_morphospace, metric = c(sum, variances))

## Disparity for the rarefied data
rare_disparity <- dispRity(rare_bin_morphospace, metric = c(sum, variances))

## Again, these objects are both from the "dispRity" class and just display
## some information when called, note the results!
boot_disparity
rare_disparity

## To display the actual results we can use the summary function
summary(boot_disparity)
summary(rare_disparity)
```

@@

```{r}
## See more options on the summary function:
?summary.dispRity
```

@@

```{r}
## Graphical options
quartz(width = 10, height = 5) ; par(mfrow = (c(1,2)), bty = "n")

## Plotting the bootstrapped and rarefied results
plot(boot_disparity, type = "continuous", main = "bootstrapped results")
plot(rare_disparity, type = "continuous", main = "rarefied results")
```

@@

```{r}
## testing the differences between bins in the bootstrapped dataset.
test.dispRity(boot_disparity, test = wilcox.test, comparison = "sequential",
    correction = "bonferroni")

## testing the differences between bins in the rarefied dataset.
test.dispRity(rare_disparity, test = wilcox.test, comparison = "sequential",
    correction = "bonferroni")
```
